Chronic imaging of cortical oxygen tension and blood flow after targeted vascular occlusion

### 2018-06-13 article.tex
Enabled line numbering in preparation for submission
Added stroke induction video
Response to Reviewer #1

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{tocloft}
\usepackage{array}

@@

% ----------------------------------------------------------------------------------------
% BEGIN DOCUMENT
% ----------------------------------------------------------------------------------------
\renewcommand{\cftdotsep}{\cftnodots}
\cftpagenumbersoff{figure}
\cftpagenumbersoff{table}
\begin{document}
\maketitle

% ----------------------------------------------------------------------------------------
% BEGIN ABSTRACT
% ----------------------------------------------------------------------------------------
\begin{abstract}
We present a dual-modality imaging system combining laser speckle contrast imaging and phosphorescence lifetime quenching to simultaneously map cortical blood flow and oxygen tension (\pO) in mice. Phosphorescence signal localization is achieved through the use of a digital micromirror device (DMD) that allows for selective excitation of arbitrary regions of interest. By targeting both excitation maxima of the oxygen-sensitive porphyrin Oxyphor PtG4, we are able to examine the effects of excitation wavelength and penetration depth on phosphorescence lifetime. We demonstrate the ability to measure differences in \pO\ between arteries and veins and large changes during an hyperoxic challenge. We dynamically monitor blood flow and \pO\ during DMD-targeted photothrombotic occlusion of an arteriole and highlight the presence of an ischemia-induced depolarization. Chronic tracking of the ischemic lesion over eight days revealed a rapid recovery, with the targeted vessel fully reperfusing and \pO\ returning to baseline values within five days. This system has broad applications for studying the acute and chronic pathophysiology of ischemic stroke and other vascular diseases of the brain.
\end{abstract}

@@

% Figure 5 - Photothrombosis
\begin{figure}
    \begin{center}
        \begin{tabular}{c}
            \includegraphics[width=6.25in]{Figure5.pdf}
        \end{tabular}
    \end{center}
    \caption {
        \label{fig:photothrombosis}
        Scale bars = 1 mm. (a) Speckle contrast images depicting the occlusion of a descending arteriole using DMD-targeted photothrombosis (Video 1, MPEG4, 16.9 MB). The red overlay indicates the 0.09 mm$^{2}$ region simultaneously illuminated for occlusion and \pO\ measurements. (b) Two arterioles (A1, A2), one vein (V1), and two parenchyma regions (P1, P2) were targeted for \pO\ measurements after stroke induction. (c) Relative blood flow and \pO\ within the targeted regions during and after photothrombosis. The green-shaded section indicates irradiation of the targeted arteriole. The arrow indicates the propagation of an ischemia-induced depolarization event.
    }
\end{figure}

@@

The chronic progression of the ischemia was tracked for eight days following photothrombosis. The perfusion of the occluded arteriole and broader effect on cortical flow was tracked using LSCI as shown in Fig.~\ref{fig:chronic}a. Tiled \pO\ maps acquired using both 445 nm and 637 nm excitation reveal the spatial extent of the oxygen deficit (Fig.~\ref{fig:chronic}b, c). The same five regions used during photothrombosis were targeted for chronic relative blood flow and \pO\ measurements (Fig.~\ref{fig:chronic}d-f). Relative blood flow was calculated using the pre-stroke (Day -0) measurements as baseline.

@@

The comparison of excitation wavelengths revealed significant differences in measured oxygen tension. Static \pO\ measurements under 445 nm illumination spanned a range five times larger than that of 637 nm illumination. This discrepancy was most noticeable in venous regions, with the shorter excitation wavelength resulting in \pO\ values around 50 mmHg whereas the longer wavelength resulted in values around 75 mmHg. Since depth penetration is heavily dependent upon wavelength\cite{Deng:2003kb}, the difference is likely the result of photons sampling a much larger volume of tissue under 637 nm illumination. An estimate of transmission through a surface vessel at both wavelengths can be obtained using the Beer-Lambert Law. Because scattering increases the distance traveled by photons in tissue, this represents the most conservative estimate of the effect of wavelength on transmission. Assuming hemoglobin is the primary absorber in blood plasma with a concentration of 2.3 mM \cite{Robles:2010cw} and 95\% SaO$_{2}$, the transmittance at 445 nm and 637 nm through a 100 $\mu$m arteriole is 0.9\% and 96.5\%, respectively. As excitation wavelength approaches the tissue optical window, the transmission of incident light significantly increases. The 445 nm light is almost entirely confined within a vessel of that caliber and does not extensively sample deeper microvasculature. Because the 637 nm light penetrates further into the brain, the measured \pO\ is skewed away from the vascular value and is more representative of a bulk volumetric average. This is consistent with prior Monte Carlo modeling that found fluorescence primarily originates from within surface vasculature at shorter wavelengths \cite{Davis:2011wj}.

@@

The chronic measurements revealed a severe post-stroke blood flow and oxygen deficit that recovered almost completely within five days. The spatial extent of the ischemia can be clearly seen on Days +1 and +2 in both the speckle contrast imagery and tiled \pO\ maps. The large gradient between the occluded vessel and surrounding tissue resembles the ischemic penumbra that the photothrombotic technique rarely produces \cite{Carmichael:2005gk}. By Day +5, the targeted vessel appeared to have fully reperfused and there was no evidence of hypoxia in the tiled \pO\ measurements. The speed of this recovery is faster than previously reported \cite{Schrandt:2015gu}, but can likely be explained by the smaller area targeted for photothrombosis, which resulted in a less severe ischemic lesion.

>>>>

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{tocloft}
\usepackage{array}
\usepackage{lineno}

@@

% ----------------------------------------------------------------------------------------
% BEGIN DOCUMENT
% ----------------------------------------------------------------------------------------
\renewcommand{\cftdotsep}{\cftnodots}
\cftpagenumbersoff{figure}
\cftpagenumbersoff{table}
\begin{document}
\linenumbers
\maketitle

% ----------------------------------------------------------------------------------------
% BEGIN ABSTRACT
% ----------------------------------------------------------------------------------------
\begin{abstract}
We present a dual-modality imaging system combining laser speckle contrast imaging and phosphorescence lifetime quenching to simultaneously map cortical blood flow and oxygen tension (\pO) in mice. Phosphorescence signal localization is achieved through the use of a digital micromirror device (DMD) that allows for selective excitation of arbitrary regions of interest. By targeting both excitation maxima of the oxygen-sensitive porphyrin Oxyphor PtG4, we are able to examine the effects of excitation wavelength on phosphorescence lifetime. We demonstrate the ability to measure differences in \pO\ between arteries and veins and large changes during an hyperoxic challenge. We dynamically monitor blood flow and \pO\ during DMD-targeted photothrombotic occlusion of an arteriole and highlight the presence of an ischemia-induced depolarization. Chronic tracking of the ischemic lesion over eight days revealed a rapid recovery, with the targeted vessel fully reperfusing and \pO\ returning to baseline values within five days. This system has broad applications for studying the acute and chronic pathophysiology of ischemic stroke and other vascular diseases of the brain.
\end{abstract}

@@

% Figure 5 - Photothrombosis
\begin{figure}
    \begin{center}
        \begin{tabular}{c}
            \includegraphics[width=6.25in]{Figure5.pdf}
        \end{tabular}
    \end{center}
    \caption {
        \label{fig:photothrombosis}
        Scale bars = 1 mm. (a) Speckle contrast images depicting the occlusion of a descending arteriole using DMD-targeted photothrombosis (Video 1, MPEG4, 26.9 MB). The red overlay indicates the 0.09 mm$^{2}$ region simultaneously illuminated for occlusion and \pO\ measurements. (b) Two arterioles (A1, A2), one vein (V1), and two parenchyma regions (P1, P2) were targeted for \pO\ measurements after stroke induction. (c) Relative blood flow and \pO\ within the targeted regions during and after photothrombosis. The green-shaded section indicates irradiation of the targeted arteriole. The arrow indicates the propagation of an ischemia-induced depolarization event.
    }
\end{figure}

@@

The chronic progression of the ischemia was tracked for eight days following photothrombosis. The perfusion of the occluded arteriole and broader effect on cortical flow was tracked using LSCI as shown in Fig.~\ref{fig:chronic}a. Tiled \pO\ maps acquired using both 445 nm and 637 nm excitation (Fig.~\ref{fig:chronic}b, c) reveal the spatial extent of the oxygen deficit, including a large gradient between the occluded vessel and the surrounding tissue on Days +1 and +2. The same five regions used during photothrombosis were targeted for chronic relative blood flow and \pO\ measurements (Fig.~\ref{fig:chronic}d-f). Relative blood flow was calculated using the pre-stroke (Day -0) measurements as baseline.

@@

The comparison of excitation wavelengths revealed significant differences in measured oxygen tension. Static \pO\ measurements under 445 nm illumination spanned a range five times larger than that of 637 nm illumination. This discrepancy was most noticeable in venous regions, with the shorter excitation wavelength resulting in \pO\ values around 50 mmHg whereas the longer wavelength resulted in values around 75 mmHg. Since depth penetration is heavily dependent upon wavelength\cite{Deng:2003kb}, the difference is likely the result of photons sampling a much larger volume of tissue under 637 nm illumination. An estimate of transmission through a surface vessel at both wavelengths can be obtained using the Beer-Lambert Law. Because scattering increases the distance traveled by photons in tissue, this represents the most conservative estimate of the effect of wavelength on transmission. Assuming hemoglobin is the primary absorber in blood plasma with a concentration of 2.3 mM \cite{Robles:2010cw} and 95\% SaO$_{2}$, the transmittance at 445 nm and 637 nm through a 100 $\mu$m arteriole is 0.9\% and 96.5\%, respectively. As excitation wavelength approaches the tissue optical window, the transmission of incident light significantly increases. The 445 nm light is almost entirely confined within a vessel of that caliber and does not extensively sample deeper microvasculature. Because the 637 nm light penetrates further into the brain, the measured \pO\ is skewed away from the vascular value and is more representative of a bulk volumetric average. This is consistent with prior Monte Carlo modeling that found fluorescence primarily originates from within large surface vasculature at shorter wavelengths \cite{Davis:2011wj}. As excitation wavelength increases, a larger fraction of the detected fluoresecence originates from beyond the vessel.

@@

The chronic measurements revealed a severe post-stroke blood flow and oxygen deficit that recovered almost completely within five days. The spatial extent of the ischemia can be clearly seen on Days +1 and +2 in both the speckle contrast imagery and tiled \pO\ maps in Fig.~\ref{fig:chronic}. The large gradient between the occluded vessel and surrounding tissue resembles the ischemic penumbra that the photothrombotic technique rarely produces \cite{Carmichael:2005gk}. By patterning illumination light instead of irradiating a circular region, the photothrombosis is contained within the targeted vessel, which allows the infarct to manifest downstream of the occlusion. By Day +5, the targeted vessel appeared to have fully reperfused and there was no evidence of hypoxia in the tiled \pO\ measurements. The speed of this recovery is faster than previously reported \cite{Schrandt:2015gu}, but can likely be explained by the smaller area targeted for photothrombosis, which resulted in a less severe ischemic lesion.
